% !TeX root = ../main.tex
%
\chapter{CAPACITY: Cryptographically-Enforced In-Process Capabilities for Modern ARM Architectures}\label{chap:capacity}

\emph{* The content of this chapter was published in Proceedings of the 2023 ACM SIGSAC Conference on Computer and Communications Security (CCS~'23). Association for Computing Machinery, New York, NY, USA, 874–888.}
\\

\epigraph{
	There is no fundamental reason why a program that you write for me should be able to delete my bank transactions file just because I can. Yet those are the permission rules imposed by the current commercial OSes for personal systems.
}{\textit{Norman Hardy}}

\

\

\clearpage

Modern hardware increasingly embeds security primitives that, if properly integrated, can enforce fine-grained protection within software.
Yet mainstream \glspl{os} continue to rely on process-based protection, introducing \emph{protection incoherencies}: while hardware features such as Intel PKU, ARM PA, and MTE provide security guarantees in userspace, the \gls{os} continues to mediate access at the coarse, process level.

This chapter presents \capacity, a hardware-assisted intra-process protection framework that reconciles these incoherencies by retrofitting a capability-based protection model into existing \gls{os} abstractions.
By transforming memory and file references into cryptographically verifiable capabilities, \capacity extends trust enforcement beyond the kernel boundary and establishes hardware-backed integrity across in-process domains.

The chater first explains how \capacity composes within the framework of coherent protection introduced in~\cref{chap:intro}.
The remainder of the chapter then details the design, implementation, and evaluation of \capacity, showing how principled retrofitting enables efficient and coherent in-process isolation on modern ARM systems.

\clearpage

\input{figures/bg-ipi}

\section{Toward Coherent In-process Isolation}

Despite the promise of fine-grained security and low performance overheads, modern hardware mechanisms do not integrate cleanly with existing \gls{os} abstractions.

\hdr{Kernel as a confused deputy}

The seminal paper by Connor et al.~\cite{connor2020pku} demonstrates this problem for PKU-based systems, highlighting the risk of \emph{confused-deputy attacks}~\cite{hardy_confused_1988} targeting the kernel~\cite{pku-pitfall}.
Specifically, in these system, the \gls{os} is unaware of an application's internal protection domains and therefore continues to enforce access control only at process granularity.
For instance, a domain may use the \texttt{mmap} system call to \emph{remove} \gls{pku} protection on a target page.
Even worse, system calls such as \texttt{process\_vm\_readv} and \texttt{send}, and kernel mechanisms like signals, read from userspace memory without regard for in-process domains, enabling a domain to bypass memory isolation~\cite{pku-pitfall,yang2025endokernel}.

Recent work proposes addressing this problem by introducing an in-process \emph{reference monitor} to vet resource accesses.
However, defining correct and comprehensive policies for in-process isolation is difficult.
Due to the semantic gap between \gls{os}-level protection and in-process enforcement, policies defined by such monitors are often incomplete, missing information flows on many system calls~\cite{yang2025endokernel,schrammel2022jenny}.
A principled approach to retrofitting security remains an open challenge.

I posit that the root cause of the confused-deputy problem is the \emph{incoherency of protection} introduced when attempting to retrofit hardware-assisted isolation into the existing process-level model enforced by the \gls{os}.
Most in-process mechanisms focus primarily on memory isolation, whereas the \gls{os} enforces access control across all resource types (e.g., memory \emph{and} files).
This incoherency introduces gaps in protection that can be exploited by attackers.

\hdr{Capacity: coherent in-process isolation}

\capacity adopts the principle of coherent protection (\cref{chap:intro}) by ensuring that hardware-assisted in-process isolation remains coherent—covering \emph{all} resource types.
Observing that PA and MTE on modern ARM processors exhibit characteristics of a capability-based protection model, \capacity proposes a capability-based scheme that coherently protects both userspace references (i.e., pointers) and references to system objects.
This approach enables protection of sensitive resources throughout their life cycles.

This coherency also enables efficiency.
\capacity benefits from hardware acceleration for reference authentication, achieving low overhead, while cryptographically authenticated references simplify enforcement (KISS principle~\cite{kiss}): both userspace and system resource references are authenticated with a single instruction.

\clearpage









% \pagenumbering{arabic}
\newcommand{\thename}{\capacity}
\input{sections/capacity/sslab-code}
\input{sections/capacity/terminologies}

\section{Abstract}
In-process compartmentalization and access control have been actively explored to provide in-place and efficient isolation of in-process security domains. Many works have proposed compartmentalization schemes that leverage hardware features.
Newer ARM architectures introduce Pointer Authentication (PA) and Memory Tagging Extension (MTE), adapting the reference validation model for memory safety and runtime exploit mitigation.
Despite their potential, these features are underexplored in the context of userspace program compartmentalization.

This paper presents \thename, a novel hardware-assisted intra-process access control design that embraces capability-based security principles.
\thename coherently incorporates the new hardware security features on ARM, based on the insight that the features already exhibit inherent capability characteristics.
It supports the life-cycle protection of the domain's sensitive objects \dash\ starting from their import from the file system to their place in memory.
With intra-process domains authenticated with unique PA keys, \thename transforms file descriptors and memory pointers into cryptographically-authenticated references and completely mediates reference usage with its program instrumentation framework and an efficient system call monitor.
We evaluate our \thename-enabled NGINX web server prototype and other common applications in which sensitive resources are isolated into different domains. Our evaluation shows that \thename incurs a low-performance overhead of approximately $17\%$ for the single-threaded and $13.54\%$ for the multi-threaded webserver.

