% !TeX root = ../main.tex
%
\chapter{Lessons learned and conclusion}\label{chap:discussion}

\epigraph{It's not the destination, it's the journey.}{\emph{Ralph Waldo Emerson}}

\

\


% \clearpage

%must
%
\capacity and \incognitos, the systems presented in this dissertation, address different layers of the stack but share a common theme: reestablishing coherence between newly defined protection mechanisms and the legacy models that underpin modern computing.
This chapter systhesizes the lessons learned from these systems for  \gls{os} security in the era of hardware-defined protection, and conclude.

\section{Lessons learned}

\hdr{Retrofitting demands principled abstraction design}

The systems described in this dissertation illustrate that security retrofitting is not merely an engineering exercise.
It requires rethinking how protection should remain coherent across hardware and software components through principled abstraction design.
The principle of coherent protection (\cref{chap:intro}) provides a useful guideline for constructing secure abstractions capable of bridging this gap.

In the domain of hardware-assisted in-process isolation, this principle is increasingly supported by recent work.
Endokernel~\cite{yang2025endokernel} introduces an in-process reference monitor that mediates and virtualizes all resource accesses.
Pegasus~\cite{dinglan2025pegasus} similarly merges frequently communicating processes into a single process using an in-process abstraction, amortizing the cost of inter-process communication.
These efforts effectively treat sub-process components as miniature processes, enforcing access control decisions in a manner analogous to how an \gls{os} manages process-level protection.
I expect that this principle will continue to hold for future hardware-assisted security retrofitting efforts.

\hdr{Security requires careful consideration across the stack}

Confused-deputy attacks on in-process isolation systems (\cref{chap:capacity}) and side-channel attacks on \glspl{cvm} highlight the need for a holistic threat model that spans the entire software stack.
\incognitos takes this approach to its logical extreme: it not only incorporates side-channel protection into the OS but also extends this protection across the whole system, covering both user and kernel code.

To avoid pitfalls and gaps in protection, future systems should continue to adopt such cross-layer reasoning, ensuring that policy definition and enforcement remain consistent across hardware, kernel, and user space.

\hdr{Security and compatibility can coexist through retrofitting}

Important ideas in security, although often phased out of mainstream systems in favor of compatibility, remain relevant.
Capability-based security, in particular, aligns surprisingly well with the programming model of many existing software interfaces, enabling fine-grained access control through retrofitting.

Inspired by approaches such as DroidCap~\cite{droidcap} and Capsicum~\cite{capsicum}, \capacity's cryptographic capability mechanism integrates cleanly with UNIX file descriptors, which already behave as capability-like interfaces~\cite{miller2003capability}.
Other recent systems follow similar principles.
For embedded system compartmentalization, D-Box~\cite{dbox} leverages capability-based techniques to specify access control for DMA devices.
Blindfold~\cite{li2024blindfold} demonstrates that capability-based security can also be retrofitted for confidential memory management under an untrusted \gls{os}, achieving both compatibility and simple enforcement.

The success of these systems leads me to believe that other influential security ideas, such as multi-level security~\cite{mccullough1987specifications} and information flow control/tracking~\cite{histar,denning1977certification,mohit2009complete}, can meaningfully improve the security of modern systems through retrofitting.
For instance, \emph{information flow tracking}~\cite{denning1977certification,mohit2009complete,enck2014taintdroid} could track the propagation of secrets across the stack, allowing \incognitos to efficiently randomize only those accesses involving sensitive data.


\section{Conclusion}
This dissertation examined how modern hardware reshapes the long-standing protection models established by \glspl{os} and argued that maintaining coherent protection across hardware and software is essential for secure system design.
Through \capacity and \incognitos, I demonstrated the \emph{principle of coherent protection} as a basis for effective security retrofitting:
\capacity showed that capability-based principles can be integrated with hardware acceleration to provide consistent access control over both memory and system resources, while \incognitos illustrated how the \gls{os}'s protection goals can be extended in confidential computing environments to effectively address side-channel threats.
As hardware continues to redefine system protection, \glspl{os} will increasingly need to adapt to these evolving capabilities, and the work presented in this dissertation offers initial steps toward that goal.
