\chapter{Introduction}
\pagenumbering{arabic}
\label{sec:intro}

\epigraph{
	This then is programming, both a tar pit in which many efforts have floundered and a creative activity with joys and woes all its own.
}{\textit{Frederick P. Brooks, Jr}, The Mythical Man-month}

\epigraph{
	I pray daily that more of my fellow-programmers may find the means of freeing themselves from the curse of compatibility.
}{\textit{Edsger W. Dijkstra}, The Humble Programmer}

\

\

The past century has witnessed rapid technological progress, with computing systems among its most transformative developments.
The advent of the Internet revolutionized global information exchange, while recent advances in artificial intelligence have accelerated scientific discovery at an unprecedented pace.
Underlying these breakthroughs are the infrastructures that enable computation—both software and hardware.
The \gls{os}, which mediates between them, remains a cornerstone of technological innovation.

Historically, the \gls{os} has served as the central authority for enforcing system security, providing a coherent protection model that governs users' access to resources.
However, advances in processor design have introduced new hardware-level security mechanisms, such as \glspl{tee}, that enforce protection models partially outside the \gls{os}'s control.
Consequently, system security is now jointly enforced by two distinct entities: the hardware and the \gls{os}.
When their respective protection models diverge, gaps or \emph{misalignments} may emerge, exposing attack vectors and undermining the intended guarantees of both.

This dissertation argues that hardware-assisted protection models must be systematically aligned with those established by the \gls{os}.
Building on this principle, I explore methods to \emph{retrofit} hardware security features into existing system architectures in a principled and coherent manner.
By restoring alignment between hardware and software protection, these methods enable systems to leverage the performance and security benefits of modern hardware while maintaining a unified protection model that is robust against modern threats.


\section{Motivation}

% Most \glspl{os} pursue three fundamental objectives.
% First, they must provide a set of \glspl{api} that are intuitive for application developers.
% This is achieved through \emph{abstractions} that describe high-level resource properties while concealing implementation details.
%
% Second, they must efficiently \emph{manage hardware resources}—including CPU time, memory, and storage—by allocating and reclaiming them as needed.
% Disciplines such as \emph{memory management}, \emph{file system design}, and \emph{scheduling} evolved to fulfill this objective.
%
% Third—and the focus of this thesis—the \gls{os} must ensure the \emph{security} of computation.
% This objective became essential with the advent of multiprogramming in the 1960s, when systems began supporting multiple concurrent users and programs.
% A malicious program could compromise the confidentiality, integrity, or availability of both the \gls{os} and other applications.
% Consequently, the \gls{os} must mediate all resource access to uphold these guarantees, even in the presence of potentially hostile users.

\subsection{Legacy OS's Protection Model}

\begin{table}[t]
	\begin{center}
		\begin{tabular}[c]{l|lll}
			       & File 1 & File 2 & File 3 \\
			\midrule
			User 1 & access &        &        \\
			User 2 &        & access &        \\
			User 3 &        &        & access \\
		\end{tabular}
		\caption{Example access control matrix specifying a simple protection model}\label{tab:example-acm}
	\end{center}
\end{table}

\glspl{os} enforce security by implementing a system of \emph{protection}~\cite{lampsonprotection} that defines access control policies between \emph{subjects} (e.g., users and processes) and \emph{objects} (e.g., memory, files, and resource handles).
Specifically, it determines which operations each subject may perform on a given object at any time.
For example, a simple policy might prohibit users from accessing resources owned by others.
In a system with three users and three files, this policy can be represented as an \emph{access control matrix}~\cite{lampsonprotection}, as shown in \autoref{tab:example-acm}.

\glspl{os} map their \emph{abstractions} onto the \emph{subjects} and \emph{objects} of this protection.
The \emph{process} abstraction, introduced in the 1960s~\cite{denning2022systems}, remains the foundation for representing subjects.
A process encapsulates relevant functionalities and interacts with system resources.
Commodity \glspl{os}, such as UNIX-based systems, associate processes with users and groups, enforcing protection on a per-process basis.

To represent objects, \glspl{os} employ the \emph{file} abstraction, which initially referred to a block of data on disk.
UNIX-inspired systems generalized this concept so that disk I/O, networking, and system services all appear as file operations, thereby unifying access under a consistent interface and simplifying access control.

Building on these abstractions, the \gls{os} enforces two primary security boundaries: one between the kernel and \emph{userspace} processes, and another among processes themselves.
The first boundary, maintained through hardware privilege rings, protects the \gls{os}'s components that enforce security (its \gls{tcb}) from tampering.
The second boundary underpins multiprogramming security, enforced through memory translation and mediation of all userspace resource accesses.
In terms of protection, two  \emph{protection domains}~\cite{lampsonprotection} are established: user space domains where applications run, and kernel domain where the kernel code executes.
In this thesis, the \emph{legacy protection} refers to the process- and file-centric one that underpins UNIX and POSIX-compatible \gls{os}.
This model is illustrated in \autoref{fig:legacy-prot}.


\input{figures/intro-protections}


\subsection{Limitations of legacy protection}

Although legacy protection has provided a solid foundation for decades, its assumptions no longer align with modern computing environments.
Contemporary software systems are increasingly large and complex—a peril underscored by the infamous Heartbleed vulnerability~\cite{heartbleed}, which exploited a bug in the OpenSSL cryptographic library.

Vulnerabilities in complex systems call for the \emph{separation of privilege} within complex programs into smaller, isolated modules~\cite{capacity,erim,hakc,rlbox,privtrans,gudka2015clean}.
The coarse-grained, process-level protection of traditional \glspl{os} becomes a bottleneck for this decomposition, imposing significant development and performance costs~\cite{erim,capacity}.

At the same time, emerging \emph{confidential computing paradigms}~\cite{confidential,google-confidential-computing,sgx,kaplan_amd_nodate,cca} invert traditional trust hierarchies by placing the \gls{os} ouside of the trust boundary.
These paradigms establish \emph{secure enclaves} \dash\ containers that are isolated  from the rest of the system, even the OS or the hypervisor.
Such paradigm shift introduces a completely different trust model that legacy protection were not designed to expressed.
Yet, previous systems retrofit such boundary into the legacy protection, introducing attack surfaces such as the Iago attacks \cite{checkoway2013iago} that exploits the system call APIs.


% Recent development places the entire OS inside secure enclaves to remain compatible with nnormal workloads. now the OS large code base becomes a liability for attacks.


% This thesis argues that sustaining strong security requires revisiting and aligning these legacy models with modern system realities.
% 
% [[PLACEHOLDER: specify concrete motivation for revisiting the protection model—e.g., scalability, hardware integration, or deployment feasibility.]]

\subsection{Rise of modern hardware-assisted protection}

Historically, the \gls{os} assumed complete mediation and exclusive control over hardware resources.
However, given they previously described limitations of legacy protection, modern hardware increasingly implements its own protection and isolation mechanisms, often beyond the \gls{os}'s full visibility or authority.

Taking advantages of these new features, a plenthora of systems has been developed and deployed, providing models of \emph{hardware-assisted protection}.
This thesis focuses on the two recent development in hardware-assisted protection: in-process isolation and VM-level confidential computing, or \gls{cvm}.
These developments represent to complementary direction being pushed by the industry. 

\begin{itemize}
  \item In-process isolation \emph{compresses} the protection domains inward to create \emph{sub-process isolation domains}, leveraging CPU features like Intel \gls{pku} that enable  userspace processes to establish memory isolation policies.
Works like ERIM and Hodor~\cite{erim,hodor} leverage these hardware extensions to created protected domain that isolate secret and contain the impact buggy code.
\item The \gls{cvm} paradigm \emph{expands} the protection domain outward from previous process level secure enclaves to entire VM.
This direction leverage recent advancement in TEE techonlogies introduced by Intel TDX, AMD SEV or ARM CCA to transform an entire VM, including applications and OS, into an protection domain, isolated from the untrusted hypervisor. 
Cloud
\end{itemize}



The two direction that are the interest of this thesis are


For instance, Intel \gls{pku} allows  while AMD SEV, Intel TDX, and ARM CCA introduce VM-level isolation that protects guest operating systems from malicious hypervisors.





\subsection{Problem of protection misalignment}

When applied in an ad hoc manner, these \emph{hardware-assisted protections} introduce \emph{misalignments} in protection models that can undermine system security (\autoref{tab:protection-models}).
Misalignment arises when the protection domains defined by hardware and those assumed by the operating system diverge.

For example, a hardware isolation feature may enforce access restrictions unknown to the OS, leading to inconsistencies in privilege enforcement, or vice versa.
Prior work has demonstrated \emph{confused deputy} attacks on \gls{pku}-based in-process isolation systems, arising from the \gls{os}'s lack of awareness of hardware-enforced protection domains~\cite{pku-pitfalls}.

Similarly, VM-level isolation (e.g., AMD-SEV) can exacerbate side-channel and interface vulnerabilities, as demonstrated by recent attacks against confidential virtual machines~\cite{wesee,heckler,ciphersteal,li2021cipherleaks}.
Such discrepancies can undermine both the enforceability and verifiability of system-wide security guarantees.



\section{Approach}


\subsection{Retrofitting for security}

Despite substantial advances in hardware and increasingly sophisticated attacks, OS protection has remained largely unchanged in mainstream \gls{os}, since the 1970s.
Modern applications continue to depend on long-standing standards such as \gls{posix}~\cite{posix} and the Windows API~\cite{winapi}, which constrain compatibility with alternative designs.
Consequently, current computing infrastructures still rely on legacy protection.

Due to the prevalence of the legacy protection, the practical path to introduce security has been \emph{retrofitting} \dash\ integrating security features into existing protection models.
For example, SELinux~\cite{selinux} extends Linux’s \gls{dac} model with \gls{mac} enforcement, while Capsicum~\cite{watson2010capsicum} retrofits a capability-based \gls{api} in FreeBSD to enable fine-grained access control~\cite{freebsd-capsicum}.
Similarly, large applications such as Chrome and Firefox are retrofitting sub-process isolation to contain vulnerabilities from compromised third-party libraries~\cite{rlbox}.



\begin{table}[t]
	\centering
	\small
	\begin{tabular}{p{3cm}p{3.5cm}p{2.5cm}p{4cm}}
		\toprule
		Legacy OS protection
		 & Hardware feature
		 & Hardware-defined protection
		 & Implications of protection misalignment                                                                                 \\
		\midrule
		\multirow{2}{3cm}{Inter-process isolation (files \& memory), user/kernel isolation}
		 & AMD SEV~\cite{advanced_micro_device_sev_2022}, ARM CCA~\cite{cca}, Intel TDX~\cite{tdx}
		 & VM-hypervisor isolation
		 & Amplified side-channel threats~\cite{ciphersteal,li2021cipherleaks}, interface vulnerabilities~\cite{civ,heckler,wesee} \\
		\cmidrule(lr){2-4}
		 & Intel PKU~\cite{libmpk}, ARM MTE~\cite{mte}
		 & Intra-process isolation (memory)
		 & Confused deputy attacks~\cite{pku-pitfalls}                                                                             \\
		\bottomrule
	\end{tabular}
	\caption{Misalignments between legacy OS protection and modern hardware protection}\label{tab:protection-models}
\end{table}



To fully capitalize on modern hardware features, systems must \emph{align} protection models between the legacy \gls{os} and hardware.

For in-process isolation systems, alignment requires consistent management of both file and memory access.

For VM-level enclaves, the guest \gls{os} must integrate the enclave’s threat model—including side-channel and interface attacks—into its own protection framework.


% This thesis lies at the intersection of legacy operating system design and emerging hardware security paradigms.

% It argues that while modern processors provide powerful protection primitives, their full security potential can only be realized when \glspl{os} integrate them coherently.
%
% This thesis examines two layers of the protection hierarchy to illustrate the broader principle of protection alignment.
% it argues that sustaining strong system security requires a principled *alignment* between legacy operating system protection models and emerging hardware-based isolation mechanisms.
% It explores the  through two systems: \capacity and \incognitos.


\section{Contributions}

This thesis positions itself at the intersection of legacy operating system design and emerging hardware security paradigms.
It argues that while modern processors introduce powerful protection primitives, their security potential can only be realized when operating systems evolve to integrate them coherently.
The work presented in this thesis, \capacity and \incognitos, examines two layers of the protection hierarchy to illustrate the broader principle of protection alignment



\begin{itemize}
	\item[-]
	      At the intra-process level, \capacity demonstrates how intra-process domains established through modern ARM hardware acceleration can be coherently
	      retrofitted into legacy abstractions, with a novel cryptographic capabilities mechanism.
	      \capacity retrofits capability-based protection into legacy process abstractions, leveraging ARM Pointer Authentication (\gls{pa}) and Memory Tagging (\gls{mte}) to transform resource references into cryptographically verifiable capabilities.
	      This design enforces consistent security semantics across file and memory accesses, bridging the gap between software abstractions and hardware enforcement.
	\item[-]
	      At the system level, \incognitos explores how \glspl{os} can collaborate with hardware-enforced VM-level isolation to enforce coherent threat models, achieving performant and robust.
	      \incognitos redefines the role of the operating system under hardware-enforced isolation, treating it as an active participant in the threat model of trusted execution environments.
	      By coordinating scheduling and memory layout randomization, it mitigates side-channel leakage and strengthens enclave isolation.
\end{itemize}

Together, these systems demonstrate that harmonizing OS and hardware protection models yields architectures that are both more secure and more efficient—pointing toward a new design paradigm for hardware-conscious operating system security.


\section{Outline}

\cref{chap:background} estabilshes the fundamental background

\cref{chap:capacity}

\cref{chap:incognitos}

\cref{chap:discussion} discuss

\cref{chap:conclusion}



% \begin{itemize}
% 	\item[-]
% 	      \emph{How can hardware-based in-process isolation primitives be retrofitted into legacy OS protection models in a principled manner?}
%
% 	      This question is addressed through the design and implementation of \capacity.
%
% 	      Retrofitting must be systematic to avoid the misalignments discussed earlier.
% 	      To address this, I introduce \capacity, a system that integrates capability-based security into in-process domains~\cite{cheri,capsicum,eros,cheri-jni}.
% 	      Leveraging ARM features such as Pointer Authentication (\gls{pa}) and Memory Tagging (\gls{mte}), \capacity transforms process references into cryptographically verifiable capabilities.
% 	      It enforces consistent protection across the resource lifecycle—from file import to memory access—providing hardware-backed integrity and confidentiality guarantees.
% 		      [[PLACEHOLDER: brief summary of \capacity’s evaluation results or contribution impact.]]
%
% 	\item[-]
% 	      \emph{How can OS legacy protection model be extended to align with emerging hardware protection mechanisms?}
%
% 	      this question is explored through \incognitos, which reconceptualizes the OS as a participant in enclave security
%
%
% 	      Certain hardware mechanisms (e.g., TEEs and CVMs) fundamentally alter system protection assumptions, requiring a redefinition of OS security roles.
%
% 	      I argue that the legacy \gls{os} protection model must be made aware of hardware threat models and explicitly incorporate them into its design.
%
% 	      \incognitos reimagines the \gls{os} as an active participant in shaping the side-channel resilience of \gls{tee} applications.
% 		      [[PLACEHOLDER: clarify \incognitos’ key mechanisms and results—e.g., adaptive scheduling, memory re-randomization, performance impact.]]
% \end{itemize}
%
% This thesis demonstrates that harmonizing legacy OS abstractions with modern hardware protection mechanisms enables systems that are both more secure and more efficient. Through \capacity and \incognitos, it shows that coherence between software and hardware enforcement is not merely an implementation concern but a foundational requirement for sustainable system security.
%
% % Thus, the practical path forward lies in \emph{retrofitting} \dash\ systematically integrating modern protection into existing OS protection models.
% % For example, SELinux~\cite{selinux} extends Linux’s \gls{dac} model with \gls{mac} enforcement, while Capsicum~\cite{watson2010capsicum} retrofits a capability-based \gls{api} in FreeBSD to enable fine-grained access control~\cite{freebsd-capsicum}.
% % Similarly, large applications such as Chrome and Firefox are retrofiting finer-grained isolation to contain vulnerabilities from compromised third-party libraries~\cite{rlbox}.
% %
% % Early systems such as MULTICS introduced protection domains and privilege rings—concepts still fundamental today.
% % Over the past six decades, numerous secure OS architectures (e.g., security kernels, capability-based OSes, microkernels) have been proposed, each offering stronger isolation and formal security guarantees.
% % However, few have achieved mainstream adoption.
% % Modern applications depend on long-standing standards—particularly the \gls{posix} interface—which limits compatibility with alternative architectures.
% % As a result, dominant infrastructures today rely on monolithic systems such as Linux and Windows that prioritize \gls{posix} compliance over adopting new security paradigms.
%
%
%
%
%
%
% % The \gls{os} acts as a \emph{reference monitor} mediating all system accesses.
% % Accordingly, it is assumed to be part of the trusted computing base (\gls{tcb}).
% % Interaction across these boundaries occurs solely through system calls.
% % The \gls{os} validates each request—for example, whether the process owner has the necessary permission to access a file—before granting access.


% \section{Misalignment between OS and hardware-defined protection}

